SPECS-decadal-MOS

### 2017-05-05 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r, eval=FALSE}
library(knitr)
purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
```

@@

***Gridded precipitation***

@@

```{r,warning=FALSE}
X <- retrieve('~/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
map(X,new=FALSE)
```

@@

Predictors for calibration: check for any dependency

@@

Any connection between temperature and dry spells?

```{r, fig.height=8}
ds.t2m <- DS(eof.dry,eof.t2m)
plot(ds.t2m)
```

Any connection between sea-level pressure anomalies and dry spells?

```{r, fig.height=8}
ds.slp <- DS(eof.dry,eof.slp)
plot(ds.slp)
```

>>>>

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. To strip off all the text and make a pure R-script, then use `purl`as shown below:

```{r, eval=FALSE}
## Extract just the R-code
library(knitr)
purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
```

@@

### Gridded precipitation ###

@@

```{r,warning=FALSE}
X <- retrieve('~/data/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
map(X,new=FALSE)
```

@@

Only the 4 leading EOFs for the seasonal mean dry spell length are relevant. The higher orders are noise. This is according to expectations, i.e. that droughts and heat waves are large-scale phenomena.

```{r}
eof.dry <- subset(eof.dry,ip=1:4)
```

### Predictors for calibration: check for any dependency

@@

Examine the connection between temperature and dry spells:

```{r, fig.height=8, warning=FALSE}
ds.t2m <- DS(eof.dry,eof.t2m)
plot(ds.t2m)
```

Cross-validation scores of 0.33 for the leading EOF of the mean duration of dry spells and a temperature anomaly over Iberia and souther France suggest a real link.

There is also a connection between sea-level pressure anomalies and dry spells.

```{r, fig.height=8, warning=FALSE}
ds.slp <- DS(eof.dry,eof.slp)
plot(ds.slp)
```

Cross-validation score of 0.43 and high SLP anomalies over France. The summertime temperature and SLP are expected to be related:

```{r, fig.height=8, warning=FALSE}
cca.slp.t2m <- CCA(eof.slp,eof.t2m)
plot(cca.slp.t2m)
```

The CCA suggests that there is covariance between SLP and temperature, and we need to avoid double counting of these effects.

### Residual:

```{r, fig.width=8, fig.height=8}
## Estimate the residuals
eof.res <- subset(eof.dry,it=ds.slp)
coredata(eof.res) <- coredata(eof.res) - coredata(ds.slp)
```

Check to see if the residual contains covariance with the temperature

```{r, fig.height=8, warning=FALSE}
ds.res <- DS(eof.res,eof.t2m)
plot(ds.res)
```

### Dry conditions and wild fire statistics

Use the EOFs describing the dry conditions as predictors and wild fire statistics as predictand to assess the connection

```{r}
# z <- DS(wfire,eof.dry)
```

### Projections for the future

Need to read the predictors again to get all months

```{r}
T2M <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-12,35),lat=c(27,50))
SLP <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-12,35),lat=c(27,50))
## Also - fix the index of eof - set to year.
index(eof.dry) <- year(eof.dry)
```

Two seperate analyses based on SLP and temperature respectively, assuming the RCP 4.5 emission scenario. Start with the SLP and the original data

```{r, warning=FALSE}
if (!file.exists('dse.mld.slp.rda')) {
  dse.mld.slp <- DSensemble.eof(eof.dry,predictor=SLP,it='mjjas',path="~/data/CMIP5.monthly",pattern = "psl_Amon_ens_")
  save(dse.mld.slp,file = 'dse.mld.slp.rda')
} else load('dse.mld.slp.rda')
plot(dse.mld.slp)
```

Follow up with the residuals and the large-scale temperture

```{r, warning=FALSE}
if (!file.exists('dse.mld.t2m.rda')) {
  dse.mld.t2m <- DSensemble.eof(eof.res,predictor=T2M,it='mjjas',path="~/data/CMIP5.monthly",pattern = "tas_Amon_ens_")
  save(dse.mld.t2m,file = 'dse.mld.t2m.rda')
} else load('dse.mld.t2m.rda')
plot(dse.mld.t2m)
```

Add the the two components to get an estimate of the total change. The `dse`objects are lists which contain `zoo` time series:

```{r}
dse.mld <- dse.mld.slp
gcmnm <- names(dse.mld)
gcmnm <- gcmnm[-grep('info',gcmnm)]; gcmnm <- gcmnm[-grep('eof',gcmnm)]
for (nm in gcmnm) {
  coredata(dse.mld[[nm]]) <- coredata(dse.mld.slp[[nm]]) + coredata(dse.mld.t2m[[nm]])
}
```

### 2017-05-05 EU-Circle/eu-circle_case-study-2.Rmd
EU-Circle

```{r esd}
library(esd)
```

## analysis

Wind, stormtracks, wave heights, ans sea level.

```{r pressure}
data(imilast.M03)
```

>>>>

Load the R-library ```esd``` which is used for data analysis.

```{r,message=FALSE}
library(esd)
```

## Storm and Sea Surge at a Baltic Sea Port

An oil piping transportation system is operating at Gdynia Port, Poland, one of the Baltic Oil Terminals. The system is designated to receive, store and send carriages and cars containing oil products and loading oil tankers. The purpose of the case study is to identify and predict safety risks associated with maritime transportation of petrol and oil. The transport of dangerous chemicals on sea is reasonably safe at normal environmental conditions. However, the transported goods may be swept overboard as a result of rough weather and hard sea conditions. The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of maritime accidents during the transporation of dangerous goods. Approximately one major accident happen at the Baltic Sea every year. There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment.

Here, we consider the potential stress of weather on the operation conditions, focusing specifically on maritime storms and hard sea conditions. Environmental conditions such as the sea level, wave height, wind speed and direction, and the occurence of storms affect the operation of the oil piping transporation system. Meteorological observations of these parameters are used to investigate the weather conditions at Gdynia Port and the Baltic Sea region.

**Wind**

Synoptic storms are associated with winds, storm surges and waves.

```{r}
load('cyclones.baltic.rda')
map(cyclones.baltic)
```

Plot the number of trajectories over time

```{r}
plot(cyclones.baltic)
```

***Import E-OBS data***

Observations of the wind are sparse in the Baltic region, but approximate values of wind speed and direction can be calculated from the sea level pressure (slp). A gridded slp product from E-OBS dataset is used which is made available by the EU-FP6 project ENSEMBLES (http://ensembles-eu.metoffice.com) and the data providers in the ECA&D project (http://www.ecad.eu).

Import the E-OBS slp data using the retrieve function of the esd package. The data is represented with lower-case $x$.

```{r,warning=FALSE}
if (!file.exists('eu-circle_case-study-2_slp.rda')) {
  x <- retrieve('pp_0.25deg_reg.nc',path='~/storeB/data/data.ECAD',
                lon=c(10,30),lat=c(53,66))
  save(file='eu-circle_case-study-2_slp.rda',x)
} else load('eu-circle_case-study-2_slp.rda')
## Only keep the ECAD data
map(x,FUN='q95',new=TRUE)
diagnose(x)
```

***Geostrophic wind***

Calculate the geostrophic wind based on the slp data.

**Ship accidents in the Baltic Sea**

```{r}
baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
uv <- geostrophicwind(x)
```

### 2017-05-05 EU-Circle/eu-circle_case-study-5.Rmd
EU-Circle

The calibration is fairly successful in identifying a connection between OLR and precipitation intencity. To see if other factors such as moisture has an effect, we estimate the residual from this regression analysis and repeat the analysis using saturation water vapour as predictor.

@@

>>>>

The calibration is fairly successful in identifying a connection between OLR and precipitation intencity. To see if other factors such as moisture has an effect, we estimate the residual from this regression analysis and repeat the analysis using saturation water vapour as predictor. This way, we avoid correlation/overlap between the two factors.

@@

## Projections

Need to read the predictors again to get all months

```{r}
T2M <- retrieve('~/Downloads/ERAINT-t2m-mon.nc', lon=c(-90,40),lat=c(30,70))
OLR <- retrieve('~/Downloads/ERAINT-OLR-mon.nc', lon=c(-50,40),lat=c(40,70))
## Also - fix the index of eof - set to year.
index(eof.dry) <- year(eof.dry)
```

Two seperate analyses based on OLR and saturation vapour pressure respectively, assuming the RCP 4.5 emission scenario. The first iteration uses the original predictand whereas the latter uses the residuals.

```{r, warning=FALSE}
if (!file.exists('dse.mu.olr.rda')) {
  dse.mu.olr <- DSensemble.eof(pca.mu,predictor=OLR,it='djf',path="~/data/CMIP5.monthly",pattern = "rlut_Amon_ens_")
  save(dse.mu.olr,file = 'dse.mu.olr.rda')
} else load('dse.mu.olr.rda')
plot(dse.mu.olr)
```

```{r, warning=FALSE}
if (!file.exists('dse.mu.es.rda')) {
  dse.mu.es <- DSensemble.eof(pca.res,predictor=SLP,it='djf',path="~/data/CMIP5.monthly",pattern = "tas_Amon_ens_")
  save(dse.mu.es,file = 'dse.mu.es.rda')
} else load('dse.mu.es.rda')
plot(dse.mld.slp)
```

Need to add the two together.

### 2017-05-05 SPECS/specs-decade-mos.Rmd
EU-Circle

print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/gfdl_DECA.rda')

>>>>

print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/gfdl_DECA.rda')
psl_DECA.gfdl <- psl_DECA; rm('psl_DECA'); tas_DECA.gfdl <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.gfdl$Data)
#[1]   9  10 480  41  45
print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/hadcm3_DECA.rda')
psl_DECA.hadcm3 <- psl_DECA; rm('psl_DECA'); tas_DECA.hadcm3 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.hadcm3$Data)
#[1]   6  10 480  33  30
load('~/data/miroc5_DECA.rda')
psl_DECA.miroc5 <- psl_DECA; rm('psl_DECA'); tas_DECA.miroc5 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.miroc5$Data)
#[1]   6   6 480  59  79
rm("psl_DECA_annual","tas_DECA_annual"); gc(reset=TRUE)
## Need to regrid (regrid.matrix) and combine the different forecasts into one object:
```
